
---
title: "R & Co."
author: "Helske & Helske"
date: "13. marraskuuta 2015"
output: ioslides_presentation
---
<!-- -->
## Working more efficiently with R

Kuva: Rstudio, Git, Github, Rmarkdown, Shiny, knitr, bitbucket, Rcpp, stack overflow

## R

- More and more popular (also outside academia)
- 6th most popular programming language by IEEE Spectrum
- UseR! 2015 conference: 660 participants, 284 from industry
- https://www.r-project.org/

## Rstudio

- Open source graphical interface for R and related tools
- Syntax highlighting, autocompletion, and error detection
- R projects for easy switching between jobs
- Autosave, command history, plotting history, environment browser, integrated help
- https://www.rstudio.com/

## R Markdown

- This is an R Markdown presentation.
- Simple formatting syntax for authoring HTML and PDF documents and slides
- Based on Markdown and knitr
- Closely integrated to Rstudio (New File -> R Markdown)
- http://rmarkdown.rstudio.com/

## Rmd features

- Equations like in LaTex: 
  \$y_i = \\beta x_i\$ produces $y_i = \beta x_i$       
- Embedding R code in the document:
```{r}
summary(cars)
```

## Rmd figures
```{r}
plot(cars)
```

## Git

- Version control system originally for software development
- Useful for writing research articles, PhD theses, R packages, ...
- Basics are easy and sufficient for small projects
- Embedded in Rstudio (Tools -> Version Control)
- http://rogerdudler.github.io/git-guide/

## Github

- Web-based Git repository hosting service
- Command line or graphical interface
- Bug tracking, feature requests, task management, ...
- Free for public repositories
    + Free private repositories e.g. in Bitbucket
- Easy way of interacting with other projects, developers, and users of own projects
- https://github.com/helske

## R package creation

- Sharing work with others 
    + Co-workers, CRAN etc.
- Personal projects
    + Updating functions, managing files
- Easier integration with C/C++/Fortran codes with R
- Getting started with `package.skeleton`
- In Rstudio: New Project -> New Directory -> R package


## Useful packages I

- `knitr`
    + Dynamic report generation (LaTeX, html, Markdown, ...)
- `ggplot2` and `ggvis`
    + Plotting systems for R, based on grammar of graphics
    + Easy to produce complex multi-layered graphics
- `magritter`
    + %>%
- `data.table` and `dplyr`
    + "More efficient data frame"

## Useful packages II

- `shiny`
- `test_that`
- `devtools`
- `Rcpp`
    + Package for simple C++ integration in R
    + Writing C++ separately, Rcpp handles the ugly stuff
