---
title: "R & Co."
author: "Helske & Helske"
date: "13. marraskuuta 2015"
output: ioslides_presentation
---
<!-- -->
## Working more efficiently with R

Kuva: Rstudio, Git, Github, Rmarkdown, Shiny, knitr, bitbucket, Rcpp, stack overflow

## R

- More and more popular (also outside academia)
- 6th most popular programming language by IEEE Spectrum
- UseR! 2015 conference: 660 participants, 284 from industry
- https://www.r-project.org/

## Rstudio

- Open source graphical interface for R and related tools
- Syntax highlighting, autocompletion, and error detection
- R projects for easy switching between jobs
- Autosave, command history, environment browser, integrated help
- https://www.rstudio.com/

## R Markdown

- This is an R Markdown presentation.
- R Markdown is a simple formatting syntax for authoring HTML and PDF documents and slides
- Based on Markdown and knitr
- Closely integrated to Rstudio (New File -> R Markdown)
- http://rmarkdown.rstudio.com/

## Rmd features

- Equations are (mostly) written like in LaTex: $y_i = \beta x_i$
- You can easily embed R code to document, which is run during knitting:
```{r}
summary(cars)
```

## Rmd figures
And you your favourite plots are embedded to document automatically:
```{r}
plot(cars)
```

## Git

- Version control system originally for software development
- Useful also for writing research articles, PhD thesis, R packages...
- Basics are easy, for small projects no need to use all features (correctly)
- This too is embedded to Rstudio (Tools -> Version Control)
- http://rogerdudler.github.io/git-guide/

## Github

- Web-based Git repository hosting service
- Server for remote storage of git projects with extra layer of fancy stuff
- Easy way of studying and collaborating with other projects
- https://github.com/helske

## R package creation

- Packaging is useful even for personal projects 
- Easier integration with C/C++/Fortran codes with R
- Getting started easy with `package.skeleton`
- In RStudio: New Project -> New Directory -> R package

# Rcpp

- Package for lot easier C++ integration in R
- You can (almost) write separate C++ codes and Rcpp takes care of all the ugly stuff


